---
title: "Eu 1km grid vs NL grid"
author: "Edwin de Jonge"
date: "2/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)

library(sf)
library(raster)
library(tmap)
make_grid <- function(map, res=1000){
  ext = extent(map)
  ext[1] <- floor(ext[1]/res) * res
  ext[2] <- ceiling(ext[2]/res) * res
  ext[3] <- floor(ext[3]/res) * res
  ext[4] <- ceiling(ext[4]/res) * res
  raster(map, ext=ext, res = res)
}

CRS_NL <- st_crs(28992)
CRS_EU <- st_crs(3035)

gemeente_2017 <- 
  st_read("http://cartomap.github.io/nl/wgs84/gemeente_2017.geojson") %>% # for different year change this
  st_transform(CRS_NL) # into 

dord <-
  gemeente_2017 %>% 
  dplyr::filter(statnaam %in% c("Dordrecht"))

# empty grid
grid_rd <- make_grid(dord, 1000)

# include environment of dord 
dord_eo <- st_crop(gemeente_2017, extent(grid_rd))

# chop map into 1km pieces
dord_rd <- rasterize(dord_eo, grid_rd)

# and transform the pieces back into a polygon file
dord_gridded_rd <- 
  rasterToPolygons(dord_rd) %>% 
  st_as_sf()

dord_rdgrid_eu <-st_transform(dord_gridded_rd, CRS_EU)

gemeente_eu <- gemeente_2017 %>% st_transform(CRS_EU) 

dord_eu <-
  gemeente_eu %>% 
  dplyr::filter(statnaam %in% c("Dordrecht"))

grid_eu <- make_grid(dord_eu, 1000)
dord_eu <- st_crop(gemeente_eu, extent(grid_eu)) %>% dplyr::mutate(statcode = droplevels(statcode))
g_dord_eu <- rasterize(dord_eu, grid_eu)

dord_eugrid_eu <- 
  rasterToPolygons(g_dord_eu) %>% 
  st_as_sf()

```

# EU-grid vs NL-grid

The NL-grid en EU-grid differ in the following aspects: 
- they use a different projecten (Rijksdriehoekstelsel (28892) vs Lambert Equal Area (3035))
- the 1km grid origins are aligned in the original projection, so that the boundaries of the grid cells also have nice rounded coordinates.

Lets look at a small piece of the Netherlands in the EU projection:

```{r}
background <- tm_shape(dord_eu) + tm_fill("lightgray") + 
  tm_borders("white", alpha=0.8) +
  tm_text("statnaam", alpha=0.5, size=0.7)

print(background)
```

```{r, fig.cap="EU 1km grid"}
background + tm_shape(dord_eugrid_eu) + tm_borders("gray")
```

When the NL grid is reprojected in the EU projection, one readily sees that the grids are not nicely aligned.

```{r, fig.cap="NL 1km grid in EU projection"}
background + tm_shape(dord_rdgrid_eu) + tm_borders("gray")
```


```{r, fig.cap="EU 1km grid (blue) vs NL 1km grid (red) in EU projection"}
background + 
  tm_shape(dord_eugrid_eu) + tm_borders("blue") +
  tm_shape(dord_rdgrid_eu) + tm_borders("red")
```
